A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface

### 2021-11-05 sections/healthcare_application/main.tex
Apply changes to healthcare application section

There are numerous news articles that focus on the problems that arise when 
ambulances stay parked outside of the hospital for a long amount of time
~\cite{dailyrecords}, \cite{staffordshirelive}.
Some news reports also comment on the long idle time of ambulances when 
they are not in use~\cite{herefordtimes} and there are several reports of 
examples where this became an issue for new patients~\cite{southwalesargus}.
A particular article looked into this problem from the point of view of an 
ambulance paramedic:

@@

The patients that are distributed by the EMS arrive at the hospital via an 
ambulance and are then either unloaded at the ED or stay blocked outside in 
the ambulance.
Whether or not the ambulance and the patient stay blocked is determined by 
the threshold that the given ED chooses to play.
High threshold indicates that the ED accepts patients even if it is relatively 
full, while low threshold means that the ED blocks ambulances more frequently.

@@

The strategies of the two EDs are the range of thresholds that they can choose
from and their utilities is the proportion of individuals whose time in the 
system is within a predetermined target time.
The EMS has to decide how to distribute its patients among the two EDs so that 
the weighted combination of the ambulance blocking time and the percentage of 
lost ambulances is minimised. 
As defined in \ref{sec:model_overview}, the interaction between the two EDs 
is a normal form game that is then used to inform the decision of the EMS.
Note that the formulated game here assumes that prior to making a choice the 
EMS knows the strategies that each ED is playing (figure 
\ref{fig:imperfect_info_game}). This corresponds to reacting to experienced 
delays.

@@

For this example the only Nash equilibrium of the game is achieved when both 
players choose a threshold of \( T_A = 10, T_B = 10 \).
This means that the two players' best response to each other is to only block
ambulances when they are full.

Nash equilibria is a theoretical measure which can be
inconsistent with intuitive notions about what should be the outcome of a 
game \cite{myerson1978refinements}.
Therefore it might not be the best way to describe human behaviour.
Since the work of Maynard Smith \cite{smith1986evolutionary} evolutionary game 
theory gives the tools for the emergence of stable behaviour.
One such model that allows for asymmetric payoffs, as is the case above, is 
replicator dynamics described in section 
(\ref{sec:methodology_learning_algorithms}).
Stable outcomes of this algorithm will correspond to a subset of Nash
equilibria but more importantly, will give a model of emergent behaviour.

The use of a learning algorithm allows to investigate, not only the outcome of 
the game, but also how that outcome is reached.
Consider figure \ref{fig:ard_by_itself}. 
By running asymmetric replicator dynamics on the system the behaviour that 
emerges can be observed.
It can be seen that for this particular set of parameters the strategies of 
the two hospitals converge over time. 
Both hospital 1 (row player) and hospital 2 (column player) seem to be playing 
the same strategy \(s_{10}\) which indicates that thresholds \(T_A = 10\) and 
\(T_B = 10\) are played.
What is more important in this example is how the two hospitals reached these
decisions which also highlights the importance of using a learning algorithm.
Hospital 2 is able to reach the decision in a short amount of time while 
hospital 1 takes longer and goes through numerous strategies to get there.

@@

To study the efficiency of the strategies being played a new concept is 
introduced that considers the ratio between each hospital's best achievable 
blocking time and the one that is being played.
This new concept is defined as the compartmentalised price of anarchy of the 
players of the game and is defined as \(PoA_i(\tilde{s})\), where 
\(i \in \{A, B\}\) to distinguish among the two players/hospitals where 
\(\tilde{s}\) is the strategy that is being played by player \( i \). 
The compartmentalised price of anarchy aims to measure inefficiencies in the 
model.
The \(PoA\) for the blocking time of player \(i\) for strategy \(\tilde{s}\) 
is given by:

@@

\textbf{Initial scenario:}
Using equation (\ref{eq:poa_compartmentalised}) and asymmetric replicator 
dynamics the emergent behaviour can be measured and the compartmentalised price 
of anarchy at every iteration for both players.
Figure \ref{fig:ard_original} shows the strategies that are being played and 
the values of \(PoA_A(s)\) and \(PoA_B(s)\) for all iterations of the 
learning algorithm until it reaches an evolutionary stable pair of strategies.

@@

\textbf{Incentivising players:}
From figures \ref{fig:ard_lambda_2} and \ref{fig:ard_num_of_servers} it can be
seen that we can change some parameters of the model to make it more efficient.
The approach used on figure \ref{fig:ard_penalty} is slightly different than 
the previous cases.
Once the played strategies in asymmetric replicator dynamics strategies 
converge the payoff matrices of the two are scaled in such a way so that the 
utilities of the selected strategy are penalised. This corresponds to a precise
policy change where more societally beneficial behaviours are incentivised.

Matrices \(A\) and \(B\) represent the original payoff matrices while matrices
\(\tilde{A}\) and \(\tilde{B}\) represent the incentivised payoff matrices.
It can be observed that matrix \(\tilde{A}\) is a scaled version of matrix 
\(A\) only on the row that is most frequently played and similarly matrix 
\(\tilde{B}\) of matrix \(B\) only on the column that is most frequently played
(matrix \(A\): row \(5\), matrix \(B\): column \(6\), see figure 
\ref{fig:ard_penalty}).
Note that the values of the payoff matrices only differ on the fourth digit
which is the raw utility of the strategies played.

>>>>

There are numerous news articles that focus on the problems that arise when 
ambulances stay blocked outside of the hospital for a long amount of 
time~\cite{dailyrecords, staffordshirelive}.
Some news reports also comment on the long idle time of ambulances when 
they are not in use~\cite{herefordtimes} and there are several reports of 
examples where this became an issue for new patients~\cite{southwalesargus}.
A particular article looked into this problem from the point of view of an 
ambulance paramedic:

@@

The patients that are distributed by the EMS arrive at the hospital via an 
ambulance and are then either offloaded at the ED or stay blocked outside in 
the ambulance.
Whether or not the ambulance and the patient stay blocked is determined by 
the threshold that the given ED chooses to play.
High threshold indicates that the ED accepts patients even if it is relatively 
full, while low threshold means that the ED blocks ambulances more frequently.

@@

The strategies of the two EDs are the range of thresholds that they can choose
from and their utilities is the proportion of individuals whose time in the 
system is within a predetermined target time.
The EMS has to decide how to distribute its patients among the two EDs so that 
the weighted combination of the ambulance blocking time and the percentage of 
lost ambulances is minimised. 
As defined in Section \ref{sec:model_overview}, the interaction between the two EDs 
is a normal form game that is then used to inform the decision of the EMS.
Note that the formulated game here assumes that prior to making a choice the 
EMS knows the strategies that each ED is playing (figure 
\ref{fig:imperfect_info_game}). This corresponds to reacting to experienced 
delays.

@@

For this example, the only Nash equilibrium of the game is achieved when both 
players choose a threshold of \( T_A = 10, T_B = 10 \).
This means that the two players' best response to each other is to only block
ambulances when they are full.

Nash equilibria is a theoretical measure which can be
inconsistent with intuitive notions about what should be the outcome of a 
game \cite{myerson1978refinements}.
Therefore it might not be the best way to describe human behaviour.
Since the work of Maynard Smith \cite{smith1986evolutionary}, evolutionary game 
theory gives the tools for the emergence of stable behaviour.
One such model that allows for asymmetric payoffs, as is the case above, is 
replicator dynamics described in Section 
(\ref{sec:methodology_learning_algorithms}).
Stable outcomes of this algorithm will correspond to a subset of Nash
equilibria but more importantly, will give a model of emergent behaviour.

The use of a learning algorithm allows to investigate, not only the outcome of 
the game, but also how that outcome is reached.
Consider Figure \ref{fig:ard_by_itself}. 
By running asymmetric replicator dynamics on the system the behaviour that 
emerges can be observed.
It can be seen that for this particular set of parameters the strategies of 
the two hospitals converge over time. 
Both hospital 1 (row player) and hospital 2 (column player) seem to be playing 
the same strategy \(s_{10}\) which indicates that thresholds \(T_A = 10\) and 
\(T_B = 10\) are played.
What is more important in this example is how the two hospitals reached these
decisions which also highlights the importance of using a learning algorithm.
Hospital 2 is able to reach the decision in a short amount of time while 
hospital 1 takes longer and goes through numerous strategies to get there.

@@

To study the efficiency of the strategies being played, a new concept is 
introduced that considers the ratio between each hospital's best achievable 
blocking time and the one that is being played.
This new concept is defined as the compartmentalised price of anarchy of the 
players of the game and is defined as \(PoA_i(\tilde{s})\), where 
\(i \in \{A, B\}\) to distinguish among the two players/hospitals where 
\(\tilde{s}\) is the strategy that is being played by player \( i \). 
The compartmentalised price of anarchy aims to measure inefficiencies in the 
model.
The \(PoA\) for the blocking time of player \(i\) for strategy \(\tilde{s}\) 
is given by:

@@

\textbf{Initial scenario:}
Using equation (\ref{eq:poa_compartmentalised}) and asymmetric replicator 
dynamics, the emergent behaviour can be measured and the compartmentalised price 
of anarchy at every iteration for both players.
Figure \ref{fig:ard_original} shows the strategies that are being played and 
the values of \(PoA_A(s)\) and \(PoA_B(s)\) for all iterations of the 
learning algorithm until it reaches an evolutionary stable pair of strategies.

@@

\textbf{Incentivising players:}
From Figures \ref{fig:ard_lambda_2} and \ref{fig:ard_num_of_servers} it can be
seen that we can change some parameters of the model to make it more efficient.
The approach used on Figure \ref{fig:ard_penalty} is slightly different than 
the previous cases.
Once the played strategies in asymmetric replicator dynamics strategies 
converge the payoff matrices of the two are scaled in such a way so that the 
utilities of the selected strategy are penalised. This corresponds to a precise
policy change where more societally beneficial behaviours are incentivised.

Matrices \(A\) and \(B\) represent the original payoff matrices while matrices
\(\tilde{A}\) and \(\tilde{B}\) represent the incentivised payoff matrices.
It can be observed that matrix \(\tilde{A}\) is a scaled version of matrix 
\(A\) only on the row that is most frequently played and similarly matrix 
\(\tilde{B}\) of matrix \(B\) only on the column that is most frequently played
(matrix \(A\): row \(5\), matrix \(B\): column \(6\), see Figure 
\ref{fig:ard_penalty}).
Note that the values of the payoff matrices only differ on the fourth digit
which is the raw utility of the strategies played.
